### 2017-06-18 article/PlosOne/S1_appendix.tex
PlosOne style manuscript (including figures) done

\documentclass{article}
\usepackage[dvips]{graphicx}

\usepackage{fullpage}
\usepackage{color}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage[table]{xcolor}
\definecolor{lightgray}{gray}{0.9}
\usepackage[T1]{fontenc}

\usepackage{caption,setspace}
\captionsetup{font={footnotesize,stretch=1.5}}
%\renewcommand{\baselinestretch}{2}

\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes, positioning, calc}

\newcommand{\bfU}{\mbox{\boldmath$\mathsf{U}$}}
\newcommand{\bfu}{\mbox{\boldmath$\mathsf{u}$}}
\newcommand{\hl}[1]{\textcolor{magenta}{#1}}
\newcommand{\RR}{\mathbb{R}}
\DeclareMathOperator*{\V}{V}
\DeclareMathOperator*{\argmax}{argmax}
\newcommand{\R}[1]{\texttt{#1}}
\newcommand{\acos}{\text{arccos}}

\begin{document}

\section*{S1 Appendix: Details about the simulated data}
This supplementary appendix describes the origin of the simulated data used for presenting the PCADSC plots. Dataset A consisted of 1000 independent simulated realizations from the $N(0, \Sigma_1)$-distribution, while dataset B consisted of 500 independent realizations from $N(0, \Sigma_1)$ and 500 independent realizations from $N(0, \Sigma_2)$. Dataset A was furthermore appended with a grouping variable, randomly dividing the observations into two groups. Dataset B also included a grouping variable, and this variable contained information about which of the two normal distributions each observation was simulated from. The covariance matrices, $\Sigma_1$ and $\Sigma_2$, were defined by

$$\Sigma_1 = \begin{pmatrix}
    1.0  &  0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
 0.0  &  1.0 & 0.0 & 0.0&  0.0 & 0.0 \\
 0.0   & 0.0 &  1.0 & 0.7 & 0.0 & 0.0 \\
 0.0 &  0.0 & 0.7 & 1.0 & 0.0 & 0.0 \\
0.0 &    0.0 &  0.0 & 0.0 & 1.0 & 0.4 \\
 0.0 & 0.0 &  0.0 & 0.0 & 0.4 & 1.0
\end{pmatrix}$$
and
$$
\Sigma_2 = \begin{pmatrix}
 1.0 & 0.2 & 0.1 & 0.0 & 0.0 & 0.0 \\
 0.2 & 1.0 & 0.1 & 0.0 & 0.0 & 0.0 \\
 0.1 & 0.1 & 1.0 & 0.0 & 0.0 & 0.0 \\
 0.0 & 0.0 & 0.0 & 1.0 & 0.3 & 0.1 \\
 0.0 & 0.0 & 0.0 & 0.3 & 1.0 & 0.2 \\
 0.0 & 0.0 & 0.0 & 0.1 & 0.2 & 1.0
\end{pmatrix},
$$
respectively.
\end{document}