### 2018-03-13 article/Old versions/MPS/titlepage.tex

\documentclass[titlepage,12pt,twoside]{article}
\usepackage[dvips]{graphicx}

\usepackage[myheadings]{fullpage}

\usepackage{authblk}
\usepackage{color}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{hyperref}
%\usepackage{colortbl}
\usepackage[T1]{fontenc}

\usepackage{caption,setspace}

\newcommand{\bfU}{\mbox{\boldmath$\mathsf{U}$}}
\newcommand{\bfu}{\mbox{\boldmath$\mathsf{u}$}}
\newcommand{\hl}[1]{\textcolor{magenta}{#1}}
\newcommand{\RR}{\mathbb{R}}
\DeclareMathOperator*{\V}{V}
\DeclareMathOperator*{\argmax}{argmax}
\newcommand{\R}[1]{\texttt{#1}}
\newcommand{\acos}{\text{arccos}}

%\begin{figure}[h]
%\centerline{\includegraphics{figure03.eps}}
%\caption{Projection of item discrimination vectors onto $V_{\theta_T}$ hyperplance for a six item three-dimensional approximate sample structure.}
%\end{figure}

%\raggedbottom
\flushbottom

%\firstpage{1}
%\setcounter{lastpage}{999}
\setcounter{secnumdepth}{3}

\title{Exploratory data structure comparisons: Three new visual tools based on Principal Component Analysis}
%\shorttitle{PCA-based data comparisons}

\author[1]{Anne Helby Petersen}
\affil[1]{Department of Public Health, University of Copenhagen}
\affil{\href{mailto:ahpe@sund.ku.dk}{\nolinkurl{ahpe@sund.ku.dk}} (Corresponding author)}

\author[2]{Bo Markussen}
\affil[2]{Department of Mathematical Sciences, University of Copenhagen}

\author[2]{Karl Bang Christensen}
\vspace{\fill}

\setlength{\parindent}{0pt}

\begin{document}

\thispagestyle{empty}

\centering

{\huge{Exploratory data structure comparisons: Three new visual tools based on Principal Component Analysis}}

\bigskip 
Short title: PCA-based data comparisons
\bigskip\bigskip

Anne Helby Petersen \\
\href{mailto:ahpe@sund.ku.dk}{\nolinkurl{ahpe@sund.ku.dk}} (Corresponding author)\\
Department of Public Health, University of Copenhagen

\bigskip

Bo Markussen \\
\href{mailto:bomar@math.ku.dk}{\nolinkurl{bomar@math.ku.dk}} \\
Department of Mathematical Sciences, University of Copenhagen

\bigskip

Karl Bang Christensen \\
\href{mailto:kach@sund.ku.dk}{\nolinkurl{kach@sund.ku.dk}} \\
Department of Public Health, University of Copenhagen

\end{document}