### 2019-12-28 dataMaid_test.Rmd

---
dataMaid: yes
title: test
subtitle: "Autogenerated data summary from dataMaid"
date: 2018-05-18 12:11:11
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
  - \newcommand{\emini}{\end{minipage}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
```{r visualFunctions, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
ggAggHist <- getFromNamespace("ggAggHist", "dataMaid")
ggAggBarplot <- getFromNamespace("ggAggBarplot", "dataMaid")
```
# Data report overview
The dataset examined has the following dimensions:

---------------------------------
Feature                    Result
------------------------ --------
Number of observations        500

Number of variables             1
---------------------------------

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

----------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character    factor    labelled   numeric    integer    logical     Date   
----------------------------------------------------- ----------- ---------- ---------- ---------- ---------- --------- ----------
Identify miscoded missing values                       $\times$    $\times$   $\times$   $\times$   $\times$             $\times$

Identify prefixed and suffixed whitespace              $\times$    $\times$   $\times$

Identify levels with < 6 obs.                          $\times$    $\times$   $\times$

Identify case issues                                   $\times$    $\times$   $\times$

Identify misclassified numeric or integer variables    $\times$    $\times$   $\times$

Identify outliers (Turkish Boxplot style)                                                $\times$   $\times$

Identify outliers                                                                                                        $\times$ 
----------------------------------------------------------------------------------------------------------------------------------

Please note that all numerical values in the following have been rounded to 2 decimals.

# Summary table

----------------------------------------------------------------------------------
&nbsp;   Variable class     # unique values   Missing observations  Any problems? 
-------- ---------------- ----------------- ---------------------- ---------------
[x]      numeric                        415                 0.20 %    $\times$    
----------------------------------------------------------------------------------

# Variable list
## x

\bminione

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                   numeric

Number of missing obs.        1 (0.2 %)

Number of unique values             414

Median                            10.22

1st and 3rd quartiles       9.91; 10.47

Min. and max.                  -Inf; 14
---------------------------------------

\emini
\bminitwo
```{r Var-1-x, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(factorV = structure(1:20, .Label = c("[3.17,3.71]", 
"(3.71,4.25]", "(4.25,4.79]", "(4.79,5.34]", "(5.34,5.88]", "(5.88,6.42]", 
"(6.42,6.96]", "(6.96,7.5]", "(7.5,8.04]", "(8.04,8.58]", "(8.58,9.12]", 
"(9.12,9.67]", "(9.67,10.2]", "(10.2,10.7]", "(10.7,11.3]", "(11.3,11.8]", 
"(11.8,12.4]", "(12.4,12.9]", "(12.9,13.5]", "(13.5,14]"), class = "factor"), 
    Freq = c(1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 4L, 58L, 
    184L, 190L, 35L, 16L, 7L, 2L, 0L, 1L), xmin = c(3.16992500144231, 
    3.71127890213383, 4.25263280282536, 4.79398670351688, 5.33534060420841, 
    5.87669450489993, 6.41804840559146, 6.95940230628298, 7.50075620697451, 
    8.04211010766603, 8.58346400835755, 9.12481790904908, 9.6661718097406, 
    10.2075257104321, 10.7488796111237, 11.2902335118152, 11.8315874125067, 
    12.3729413131982, 12.9142952138897, 13.4556491145813), xmax = c(3.71127890213383, 
    4.25263280282536, 4.79398670351688, 5.33534060420841, 5.87669450489993, 
    6.41804840559146, 6.95940230628298, 7.50075620697451, 8.04211010766603, 
    8.58346400835755, 9.12481790904908, 9.6661718097406, 10.2075257104321, 
    10.7488796111237, 11.2902335118152, 11.8315874125067, 12.3729413131982, 
    12.9142952138897, 13.4556491145813, 13.9970030152728), ymin = c(0, 
    0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), 
    ymax = c(1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 4L, 58L, 
    184L, 190L, 35L, 16L, 7L, 2L, 0L, 1L)), .Names = c("factorV", 
"Freq", "xmin", "xmax", "ymin", "ymax"), row.names = c(NA, -20L
), class = "data.frame"), vnam = "x")
```
\emini

- The following suspected missing value codes enter as regular values: \"-Inf\".

- Note that a check function found the following problematic values: \"-Inf\", \"3.17\", \"8.84\", \"8.94\", \"9.02\", \"11.35\", \"11.4\", \"11.41\", \"11.41\", \"11.47\" (21 additional values omitted).

\fullline

Report generation information:

 *  Created by Anne Helby Petersen (username: `zms499`).

 *  Report creation time: fr maj 18 2018 12:11:15

 *  Report Was run from directory: `P:/PCADSC/R`

 *  dataMaid v1.1.2 [Pkg: 2018-05-03 from CRAN (R 3.4.4)]

 *  R version 3.4.2 (2017-09-28).

 *  Platform: x86_64-w64-mingw32/x64 (64-bit)(Windows 7 x64 (build 7601) Service Pack 1).

 *  Function call: `makeDataReport(data = test, checks = setChecks(numeric = defaultNumericChecks(add = "identifyOutliersTBStyle", 
    remove = "identifyOutliers"), integer = defaultIntegerChecks(add = "identifyOutliersTBStyle", 
    remove = "identifyOutliers")))`